News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-03-24 rupnik2015.tex
update to section 3

The paper is organized as follows: we first provide an overview of the system as a whole (Section ~\ref{sec:pipeline}), followed by details on each of the two steps described above. We first introduce several approaches for measuring cross-lingual similarity (Section~\ref{sec:crosslingual}) and then the algorithms for matching clusters in different languages (Section~\ref{sec:linking}).  To improve readability, we describe relevant related work for each module in its corresponding Section. Finally, we present and interpret the experimental results, give conclusions and describe promising future directions.

@@

The processing is done independently for each language and the output is a clustering of articles in each language, with each cluster ideally representing an event, which are defined by topically, geographically and temporally. The goal of this paper is present approaches to  connecting the clusters/events across languages.

Our approach is based on first defining a similarity function between articles in different languages (Section~\ref{sec:crosslingual}) and then we present several efficient approaches to linking the clusters based on features (Section~\ref{sec:linking}). The pipeline is shown in Figure~\ref{fig:pipeline} First, however we describe the pre-processing on the input articles which we then take as input.

The system is designed to handle streams and so we describe how a document which enters the stream is processed. The sequence of operations is shown in  Figure~\ref{fig:erpipeline}.  A new document is first tokenized, stop words are removed and words are stemmed.
 What remains is represented in a vector-space model and normalized using $TF-IDF$ (see Section~\ref{sec:tfidf} for the definition). Since each article is tagged with its language, cosine similiarity  is used to find the modst similar existing cluster, by comparing it to the centroid vector of each cluster. A selected threshold is used to determine if the article is not similar to any existing clusters (0.4 was used in our experiments).  The article is then assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article.

@@

The clusters, we consider must have a sufficient number of articles (which is a language dependent paramater). This is to ensure that we consider events. Once an event is identified(i.e. a cluster of sufficient size), a new unique ID is assigned to it and the main information about the event is then automatically extracted by analyzing the articles assigned to it. The extracted information includes properties such as the date of the event, the location, who is involved in it, what the event is about, etc.

 Further details about the system can be found in \cite{}.

@@

\section{Cross-lingual Document Similarity}
\label{sec:crosslingual}
Document similarity is an important component in techniques from text mining and natural language processing. Many techniques use the similarity as a black box, i.e., a kernel in Support Vector Machines. Comparison of documents (or other types of text snippets) is a well studied problem \todo{some references to prove that}. In this section we define document similarity in a cross-lingual setting, where the similarity function receives documents in different languages. We conclude the section by an introduction of \todo{two datasets} which we used in this paper to learn cross-lingual similarity functions.

@@

\section{Multilingual Latent Factor Models}\label{sec:models}
In this section, we describe several approaches to computing multilingual simliarities described in the previous section. We present three approaches
\begin{enumerate}
\item  a simple approach based on $k$-means clustering
\item  a standard approach based on singular value decomposition
\item a new method that is applicable in more general settings.
\end{enumerate}
 We concentrate on approaches that are based on linear maps rather than alternatives, such as machine translation and probabilistic models. Applicability of machine translation relies
on the availability of training resources (strongly aligned translation corpora), which are not available for many world languages, but there exist comparable corpora. Probabilistic
methods \todo{provide reasons for not using them: e.g. difficulty with parameter estimation, maybe just say future work?}.

@@

\section{Cross-lingual Event Linking}

>>>>

The paper is organized as follows: we first provide an overview of the system as a whole (Section ~\ref{sec:pipeline}), followed by details on each of the two steps described above. We first introduce the problem of cross-lingual similarity computation (Section~\ref{sec:crosslingual}), which is followed by several approaches for measuring cross-lingual similarity (Section~\ref{sec:models}) and then the algorithms for matching clusters in different languages (Section~\ref{sec:linking}).  To improve readability, we describe relevant related work for each module in its corresponding Section. Finally, we present and interpret the experimental results, give conclusions and describe promising future directions.

@@

The processing is done independently for each language and the output is a clustering of articles in each language, with each cluster ideally representing an event, which are defined  topically, geographically and temporally. The goal of this paper is to present approaches to connecting the clusters/events across languages.

Our approach is based on first defining a similarity function between articles in different languages (Section~\ref{sec:models}) and then we present several efficient approaches to linking the clusters based on features (Section~\ref{sec:linking}). The pipeline is shown in Figure~\ref{fig:pipeline}. First, however we describe the pre-processing on the input articles which we then take as input.

The system is designed to handle streams and so we describe how a document which enters the stream is processed. The sequence of operations is shown in  Figure~\ref{fig:erpipeline}.  A new document is first tokenized, stop words are removed and words are stemmed.
 What remains is represented in a vector-space model and normalized using $TF-IDF$ (see Section~\ref{sec:tfidf} for the definition). Since each article is tagged with its language, cosine similarity  is used to find the most similar existing cluster, by comparing it to the centroid vector of each cluster. A selected threshold is used to determine if the article is not similar to any existing clusters (0.4 was used in our experiments).  The article is then assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article.

@@

The clusters, we consider must have a sufficient number of articles (which is a language dependent parameter), which is in line with our notion of an event. Once an event is identified (i.e. a cluster of sufficient size), a new unique ID is assigned to it and the main information about the event is then automatically extracted by analyzing the articles assigned to it. The extracted information includes properties such as the date of the event, the location, who is involved in it, what the event is about, etc.

 The system is described in more detail elsewhere \shortcite{Leban2014W,Leban2014I}.

@@

\section{Cross-lingual Document Similarity}\label{sec:crosslingual}
\label{sec:crosslingual}
Document similarity is an important component in techniques from text mining and natural language processing. Many techniques use the similarity as a black box, i.e., a kernel in Support Vector Machines. Comparison of documents (or other types of text snippets) in a monolingual setting is a well studied problem in the field of information retrieval \cite{Salton88term-weightingapproaches}. In this section we will first cover the related work on cross-lingual similarity models and then formally introduce the problem.

@@

\section{Cross-Lingual Models}\label{sec:models}
In this section, we describe several approaches to computing multilingual simliarities described in the previous section. We present three approaches
\begin{enumerate}
\item  a simple approach based on $k$-means clustering
\item  a standard approach based on singular value decomposition
\item a new method that is applicable in more general settings.
\end{enumerate}
 We concentrate on approaches that are based on linear maps rather than alternatives, such as machine translation and probabilistic models. Applicability of machine translation relies
on the availability of training resources (strongly aligned translation corpora), which are not available for many world languages, but there exist comparable corpora. Probabilistic
methods \todo{provide reasons for not using them: e.g. difficulty with parameter estimation, maybe just say future work?}.

@@

\section{Cross-lingual Event Linking}\label{sec:linking}
